---
title: "CpG islands in prom"
output:
  html_document:
    df_print: paged
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

To better analyze the DMR I will take into account not the promoter region but only the CpGs inside the promoters.I start the analysis with the significant regions MEDIPs object

```{r, eval=FALSE}
 setwd("C:/Users/oncotraslacional/Documents/Arantxa/MBD-Seq_David")
 #load the regions

 cpgs <- read.table(file="C:/Users/oncotraslacional/Documents/Arantxa/MBD-Seq_David/RData/intersect_CpG_with_promotors_pcgenes_-2000_TSS_+500.bed",
                    stringsAsFactors = FALSE)
 
 colnames(cpgs) <- c("chr", "start", "end", "gene", "strand")
 
 rois <- cpgs[,1:4]
 
 #Use the selectROIS function from MEDIPs
 
 cols.to.keep <- c(4:10, 17:20)
 
 clones <- names( mr.edgeR.sig)
 
 gained.cpgs <- list()
 
 lost.cpgs <- list()
 
 DMR <- list()

 for(i in 1:length(clones)){
 
   DMR[[i]] <-  MEDIPS.selectROIs( results = mr.edgeR.sig[[i]], # Keep CpGs only
                                                rois = rois,
                                                columns= colnames(mr.edgeR.sig[[i]])[cols.to.keep] )
   clone <- DMR[[i]]
  
   gained.cpgs[[i]] <- clone[ clone$edgeR.logFC > 2, ] #select gained regions
   
   lost.cpgs[[i]] <-  clone [ clone$edgeR.logFC < -2, ] #Select lost regions

}

 #Check if we have any case in which a gene is both hyper and hypomethylated
 
 dobles <- c()
 
 for( i in 1:length(clones)){
  dobles[i]<- sum(unique(lost.cpgs[[i]]$ROI) %in% unique(gained.cpgs[[i]]$ROIS))
} 

 print(dobles)
 
```
save the tables


```{r, eval=FALSE}

file.names <- paste( clones, "gained_cpgs.txt", sep="_")

for (i in 1:length(clones)){
  
  write.table( gained.cpgs[[i]], 
               file= file.names[i],
               row.names = FALSE)
}

file.names <- paste( clones, "lost_cpgs.txt", sep="_")

for (i in 1:length(clones)){
  
  write.table( lost.cpgs[[i]], 
               file= file.names[i],
               row.names = FALSE)
}


```

In the previous list we have the differential methylated CpGs inside promoters, but there might be more than one per promoter and it is also interesting to check if both or only one is a DMR. For that I will name each of the island belonging to a certain promoter gene as 1, 2, 3, etc.. and then perform the selection and merge of the regions

```{r, eval=FALSE}

regiones <- rois$gene
repetidas <- duplicated(regiones)

regiones[repetidas] <- paste(regiones[repetidas],
                             "1",
                             sep =".")

#In case there are more than 2 CpGs per gene I repeat the process

repetidas <- duplicated(regiones)

repes <- strsplit(regiones[repetidas],
                             split =".",
                             fixed = TRUE)
to.change <- c()

for(i in 1:length(repes)){
  
  to.change[i] <- paste(repes[[i]][1],
                        "2",
                        sep = ".")
}

regiones[repetidas] <- to.change #replace them


#In case there are more than 3 CpGs per gene I repeat the process

repetidas <- duplicated(regiones)

repes <- strsplit(regiones[repetidas],
                  split =".",
                  fixed = TRUE)
to.change <- c()

for(i in 1:length(repes)){
  
  to.change[i] <- paste(repes[[i]][1],
                        "3",
                        sep = ".")
}

regiones[repetidas] <- to.change #replace them


#Now change the gene column by the regiones object

rois$gene<- regiones



```
I can select now those regions from the MEDIPs object

```{r, eval=FALSE}
cols.to.keep <- c(4:10, 17:20)
 
 clones <- c("H827_1",
             "H827_12",
             "H827_4",
             "H827_6",
             "H827_pl1")
 
 gained.cpgs <- list()
 
 lost.cpgs <- list()
 
 DMR <- list()

 for(i in 1:length(clones)){
 
   DMR[[i]] <-  MEDIPS.selectROIs( results = mr.edgeR.sig[[i]], # Keep promoters only
                                                rois = rois,
                                                columns= colnames(mr.edgeR.sig[[i]])[cols.to.keep] )
   clone <- DMR[[i]]
  
   gained.cpgs[[i]] <- clone[ clone$edgeR.logFC > 2, ] #select gained regions
   
   lost.cpgs[[i]] <-  clone [ clone$edgeR.logFC < -2, ] #Select lost regions

}


 
```
The value of MEDIPs contains the significantly enriched or depleted methylated regions inside the CpG islands in 100 bp windows. I will merge adjacent windows using the function merge.by.gene.
```{r, eval=FALSE}

#Clone "H827_1"doesn't have any gained regions, so I remove it from the for loop

source("C:/Users/oncotraslacional/Documents/Arantxa/MBD-Seq_David/functions/merge_results_by_gene_function.R")


 DMR.by.gene<- vector( "list", length(clones))


for (i in 1:length(clones)){
  
 DMR.by.gene[[i]] <- merge.by.gene (DMR[[i]])
 
}


#Merge the lost regions as well
clones <- c("H827_1",
            "H827_12",
            "H827_4",
            "H827_6",
            "H827_pl1")

lost.cpgs.merge <- list()
for (i in 1:length(clones)){
  
  lost.cpgs.merge[[i]] <- merge.by.gene (lost.cpgs[[i]])
}


 


```
Save the tables as .txt
```{r, eval=FALSE}

clones <- c("H827_12",
            "H827_4",
            "H827_6",
            "H827_pl1")

file.names <- paste( clones, "gained_cpgs_merge_by_island.txt", sep="_")

for (i in 1:length(clones)){
  
  write.table( gained.cpgs.merge[[i]], 
               file = file.names[i],
               row.names = FALSE)
}

#Do the same for the lost regions

clones <- c("H827_1",
            "H827_12",
            "H827_4",
            "H827_6",
            "H827_pl1")

file.names <- paste( clones, "lost_cpgs_merge_by_island.txt", sep="_")

for (i in 1:length(clones)){
  
  write.table( lost.cpgs.merge[[i]], 
               file = file.names[i],
               row.names = FALSE)
}


```
To make the heatmaps I need to make a matrix per clon with the common genes

```{r}

to.heatmap <- vector("list", length(DMR.by.gene))
names(to.heatmap) <- names( DMR.by.gene)
  
  for (i in 1:length(DMR.by.gene)){
    
   
    to.heatmap[[i]] <- subset.data.frame(DMR.by.gene[[i]], 
                                         subset = abs(DMR.by.gene[[i]]$edgeR.logFC) > 2,
                                         select = c(gene, edgeR.logFC,edgeR ))
    
    names(to.heatmap[[i]])[2] <- names(to.heatmap)[i] 
  
  }

  #Make a data.frame per line 
   h1975 <- Reduce(function(x, y) merge(x, y, all=FALSE), to.heatmap[6:10])
   rownames(h1975)<- h1975$gene
   pheatmap( h1975[,-1], treeheight_row = 2, treeheight_col =1 )
   pc9 <- Reduce(function(x, y) merge(x, y, all=FALSE), to.heatmap[12:15])
   rownames(pc9) <- pc9$gene
   pheatmap(pc9[,-1], fontsize_row = 1, treeheight_row = 2, treeheight_col =1 )
   h827<- Reduce(function(x, y) merge(x, y, all=FALSE), to.heatmap[1:5])
   rownames(h827)<- h827$gene
   pheatmap(h827[,-1], treeheight_row = 2, treeheight_col =1 )
   
   
   
```


